# Computing - Chapter 03

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library Installation

```{r}
#install.packages("AppliedPredictiveModeling")
#install.packages("e1071")
#install.packages("caret")
#install.packages("corrplot")
```

## Load Library

```{r}
library(AppliedPredictiveModeling)
library(e1071)
library(caret)
library(corrplot)
```


## Load Data

```{r}
data(segmentationOriginal)

# look at just training data
segData <- subset(segmentationOriginal, Case == "Train")

# extract some vectors from the main dataset
cellID <- segData$Cell
class <- segData$Class
case <- segData$Case

# remove those variables 
segData <- segData[, -(1:3)]

# remove the "status" columns
statusColNum <- grep("Status", names(segData))
segData <- segData[, -statusColNum]
```

## Transformations

Determine the skewness level of the variables

```{r}
skewValues <- apply(segData, 2, skewness)
head(skewValues)
```

Visualize those needing review

```{r, echo=FALSE}
hist(segData$AreaCh1)
# OR
# library(lattice)
# histogram(segData$AreaCh1)
```

Define a box-cox transformation 

```{r}
Ch1AreaTrans <- BoxCoxTrans(segData$AreaCh1)
Ch1AreaTrans
```

Compare the original to the transformed

```{r}
# original
head(segData$AreaCh1)

# transformed
predict(Ch1AreaTrans, head(segData$AreaCh1))

# manual confirmation
(segData$AreaCh1[1] ^ (Ch1AreaTrans$lambda) - 1) / (Ch1AreaTrans$lambda)
```

Center and scale, then perform PCA

```{r}
pcaObject <- prcomp(segData, center=TRUE, scale.=TRUE)
percentVariance <- pcaObject$sd ^ 2 / sum(pcaObject$sd ^ 2) * 100
percentVariance[1:3]
```


Review the tranformed values (stored in value `x`)

```{r}
head(pcaObject$x[, 1:5])
```

Review the loadings

```{r}
head(pcaObject$rotation[, 1:3])
```

`caret` package supports `spatialSign` function: `spatialSign(segData)`

For imputing missing values, consider the `impute` package.  e.g. `impute.knn`.  (`preProcess` in the `caret` package makes use of this.)

`preProcess` can be used to generate and object to perform a series of data preparation steps.

```{r}
trans <- preProcess(segData,
                    method = c("BoxCox", "center", "scale", "pca"))
trans
```

The object generated can then be used to perform the transformation:

```{r}
transformed <- predict(trans, segData)
head(transformed[, 1:5])
```

Order of operations: transformation, centering, scaling, imputation, feature extraction, spatial sign.

The `train` function in `caret` can optionally use `preProcess` within resampling iterations.

## Filtering

`nearZeroVar` (`caret`) - returns the column numbers of any near-zero variance predictors

`cor` (`stats`) - generate a correlation matrix

```{r}
correlations <- cor(segData)
dim(correlations)
correlations[1:4, 1:4]
```

`corrplot` (`corrplot`) - visualize a correlation matrix to highlight relationships with high positive or negative correlation

```{r}
corrplot(correlations, order="hclust")
```

`findCorrelation` (`caret`) - returns column numbers for the predictors recommended for removal due to high correlation

```{r}
highCorr <- findCorrelation(correlations, cutoff = 0.75)
length(highCorr)
head(highCorr)
filteredSegData <- segData[, -highCorr]
```

The `subselect` package contains several functions that can be used to perform the same filtering.

## Creating Dummy Variables

* Recommend using the full set of dummy variables when working with tree based models

Example that attempts to predict iris sepal length based on species

```{r}
data(iris)
levels(iris$Species)

simpleMod <- dummyVars(~Sepal.Length + Species, data=iris, levelsOnly=TRUE)
simpleMod
```

Generate the dummy variables for the training set

```{r}
predict(simpleMod, head(iris))
```

Or, if we believe there's a _joint effect_ between predictors, we can include an interaction:

```{r}
withInteraction <- dummyVars(~Sepal.Length + Species + Sepal.Width:Species,
                             data = iris,
                             levelsOnly=TRUE)
withInteraction
```

Generate the dummy variables (with interactions) for the training set:

```{r}
predict(withInteraction, head(iris))
```